A sustainable approach to tourist signage on heritage trails\footnote{This research has been supported by the Ministry of University and Research under the "Underlandscape" project (2020428LS8)}

### 2023-11-04 paper.tex
Sincronizzato con - Ancora modifiche augusto

The problem with the above fact is that we humans are not able to understand the long-term consequences of deploying such a potential, nor even to keep it under control. Even worse, such capacities are often used with the aim of short-term local effects, ignoring long-term global ones, with the risk of long-term global deterioration. This attitude pervades all activities, starting with the most basic ones such as the supply of food and homes, and is often influenced by economic profit.

Leisure activities play an important role in bringing profit to niche businesses. This creates a reinforcing effect that follows the negative dynamics seen above: a leisure activity that is neutral in itself (e.g., skying) can have a significant impact due to induced effects (lifts) that strengthen the success of the activity (easier to practice) bringing more investment and impact. Unlike the beneficial effects, which are usually well localized, the negative impact has a very broad spectrum that includes environmental, social, and economic sectors, as discussed by Patthey in \cite{pat08a} relating to outdoor winter sports.

@@

The promotion of cultural heritage is situated in such a complex scenario. As noted by Jorge Otero in \cite{ote22a}, local communities benefit from the promotion of heritage: conservation activity generates social growth, while tourism provides an extra income that can be reinvested in promotion. However, just as in the case of ski resorts, there is a risk of compromising the heritage itself. This interaction raises a question of sustainability also for heritage valorization initiatives, two issues that, instead,  are often considered in a virtual synergy. As stated in the document titled 'World Heritage and Sustainable Development' by UNESCO \cite{lab17a}, calling for actions that "harness the potential of World Heritage properties and heritage in general, to contribute to sustainable development".

@@

Its {\em performance} is currently difficult to evaluate. The recording of the number of times a given tag is scanned, in theory, a significant performance index, is unreliable for technical reasons. In fact, the QR tag does not have the capability to record or transmit read events, and, on the side of the smartphone, the user may be unable to visit the URL in the tag or deny such operation for security reasons. In addition, the URL can be visited by means different from the QR reading. For this reason, we do not discuss such an index.

>>>>

The problem with the above fact is that we humans are not able to understand the long-term consequences of deploying such a potential, nor even to keep it under control. Even worse, such capacities are often used with the aim of short-term local effects, ignoring long-term global ones, with the risk of long-term global deterioration. This attitude pervades all activities, starting with the most basic ones as the supply of food and homes, and is often influenced by economic profit.

Leisure activities play are relevant in bringing profit to niche businesses. This creates a reinforcing effect that follows the negative dynamics seen above: a leisure activity that is neutral in itself (e.g., skying) can have a significant impact due to induced effects (lifts) that strengthen the success of the activity (easier to practice) bringing more investment and impact. Unlike the beneficial effects, which are usually well localized, the negative impact has a very broad spectrum that includes environmental, social, and economic sectors, as discussed by Patthey in \cite{pat08a} relating to outdoor winter sports.

@@

The promotion of cultural heritage, which the Agenda 21 for Culture identifies as a fourth pillar for sustainability \cite{fourthpillar}, is situated in such a complex scenario. As noted by Jorge Otero in \cite{ote22a}, local communities benefit from the promotion of heritage: conservation activity generates social growth, while tourism provides an extra income that can be reinvested in promotion. However, just as in the case of ski resorts, there is a risk of compromising the heritage itself. This interaction raises a question of sustainability also for heritage valorization initiatives, two issues that, instead,  are often considered in a virtual synergy. As stated in the document titled 'World Heritage and Sustainable Development' by UNESCO \cite{lab17a}, calling for actions that "harness the potential of World Heritage properties and heritage in general, to contribute to sustainable development".

@@

Its {\em performance} is currently difficult to evaluate. The recording of the number of times a given tag is scanned, in theory, a significant performance index, is unreliable for technical reasons, one being that the QR tag cannot record or transmit read events. Another is that the user may be unable or unwilling to visit the URL in the tag. Finally, the URL can be visited by means different from the QR reading. For such reasons, despite we acknowledging its importance for the stakeholders, we do not discuss such an index.
